---
title: "p8105_hw5_sc5826"
author: "Shivalika Chavan"
date: "2025-10-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
source("source/bday_sim.R")

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
```{r}
bday_sim_results = 
  expand_grid(
    group_size = 2:50,
    iter = 1:10000
  ) |> 
  mutate(
    results = map_lgl(group_size, bday_sim)
  ) |> 
  group_by(group_size) |> 
  summarize(
    probability = mean(results)
  )
```


```{r}
bday_sim_results|> 
  ggplot(aes(x = group_size, y = probability)) +
  geom_point() + 
  geom_line() + 
  geom_hline(yintercept = 0.5, color = "red") + 
  xlab("Group Size") +
  ylab("Probability") + 
  labs(
    title = "Probability of 2+ people sharing a birthday based on group size"
  ) 
```

The resulting curve is S-shaped, where the probability begins to taper around a group size of 50. At a group size of 2, the probability of at least 2 people sharing a birthday is `r round(pull(filter(bday_sim_results, group_size == 2), probability), digits = 3)`. At a group size of 50, the probability of at least 2 people sharing a birthday is `r round(pull(filter(bday_sim_results, group_size == 50), probability), digits = 3)`. To have more than a 50% probability of at least 2 people sharing a birthday, the group size needs to be at least `r min(pull(filter(bday_sim_results, probability > 0.5), group_size))`. 

